### 2018-05-03 museum.tex
comments and small changes

% A museum of no-longer-needed lemmas and such (fragments of proofs of which may however become useful later)

\begin{lemma}
 If for all $u \in \val{\tau_1}{\omega,\xi}{\Gamma}$ we have
 $t[\subst{x}{u}] \in \val{\tau_2}{\omega,\xi}{\Gamma}$, then
 $\lambda x : \omega(\tau_1) . t \in
 \val{\tau_1\to\tau_2}{\omega,\xi}{\Gamma}$.
\end{lemma}

\begin{proof}
 By Lemma~\ref{lem_val_computable} the
 sets~$\val{\tau_i}{\omega,\xi}{\Gamma}$ are candidates, so
 $\val{\tau_i}{\omega,\xi}{\Gamma} \subseteq \SN$. Let
 $u \in \val{\tau_1}{\omega,\xi}{\Gamma}$. We show
 $(\lambda x : \omega(\tau_1) . t) u \in
 \val{\tau_2}{\omega,\xi}{\Gamma}$ by induction on $\nu(t) +
 \nu(u)$. Suppose $(\lambda x : \omega(\tau_1) . t) u \leadsto
 w$. There are three cases.
 \begin{itemize}
 \item $w = t[\subst{x}{u}]$. Then
   $w \in \val{\tau_2}{\omega,\xi}{\Gamma}$ by assumption.
 \item $w = (\lambda x : \omega(\tau_1) . t') u$ with
   $t \leadsto t'$. Because $\val{\tau_2}{\omega,\xi}{\Gamma}$ is a
   candidate, still
   $t'[\subst{x}{u}] \in \val{\tau_2}{\omega,\xi}{\Gamma}$. Also
   $\nu(t') < \nu(t)$, so $w \in \val{\tau_2}{\omega,\xi}{\Gamma}$
   follows from the inductive hypothesis.
 \item $w = (\lambda x : \omega(\tau_1) . t) u'$ with
   $u \leadsto u'$. Because $\val{\tau_i}{\omega,\xi}{\Gamma}$ are
   candidates, still $u' \in \val{\tau_1}{\omega,\xi}{\Gamma}$ and
   $t[\subst{x}{u'}] \in \val{\tau_2}{\omega,\xi}{\Gamma}$. Also
   $\nu(u') < \nu(u)$, so $w \in \val{\tau_2}{\omega,\xi}{\Gamma}$
   follows from the inductive hypothesis.
 \end{itemize}
 Because $(\lambda x : \omega(\tau_1) . t) u$ is neutral and
 $\val{\tau_2}{\omega,\xi}{\Gamma}$ is a candidate, this shows
 $(\lambda x : \omega(\tau_1) . t) u \in
 \val{\tau_2}{\omega,\xi}{\Gamma}$.
\end{proof}

\begin{lemma}
 If for every type~$\tau$ and every $X \in \Cb_\tau^\Gamma$ we have
 $t \in
 \val{\sigma}{\omega[\subst{\alpha}{\tau}],\xi[\subst{\alpha}{X}]}{\Gamma}$,
 then
 $\Lambda \alpha . t \in
 \val{\forall\alpha[\sigma]}{\omega,\xi}{\Gamma}$.
\end{lemma}

\begin{proof}
 Let~$\tau$ be a type and $X \in \Cb_\tau^\Gamma$. By
 Lemma~\ref{lem_val_computable} the
 set~$\val{\forall\alpha[\sigma]}{\omega,\xi}{\Gamma}$ is a
 candidate, so $t \in \SN$. By induction on~$\nu(t)$ we show
 $(\Lambda \alpha . t) \alpha \in
 \val{\sigma}{\omega[\subst{\alpha}{\tau}],\xi[\subst{\alpha}{X}]}{\Gamma}$.
 Because~$\val{\sigma}{\omega[\subst{\alpha}{\tau}],\xi[\subst{\alpha}{X}]}{\Gamma}$
 is a candidate by Lemma~\ref{lem_val_computable}, it suffices to
 show that for every~$w$ with $(\Lambda \alpha . t) \leadsto w$ we
 have
 $w \in
 \val{\sigma}{\omega[\subst{\alpha}{\tau}],\xi[\subst{\alpha}{X}]}{\Gamma}$. There
 are two cases.
 \begin{itemize}
 \item $w = (\Lambda \alpha . t') \alpha$. Then $\nu(t') < \nu(t)$
   and
   $w \in
   \val{\sigma}{\omega[\subst{\alpha}{\tau}],\xi[\subst{\alpha}{X}]}{\Gamma}$
   follows from the inductive hypothesis.
 \item $w = t$. Then
   $w \in
   \val{\sigma}{\omega[\subst{\alpha}{\tau}],\xi[\subst{\alpha}{X}]}{\Gamma}$
   by assumption.
 \end{itemize}
\end{proof}

### 2018-05-18 museum.tex
removed eta-expansion

@@
 % A museum of no-longer-needed lemmas and such (fragments of proofs of which may however become useful later)
+
+\newcommand{\arreta}{\rightarrow_\eta}
+\newcommand{\etalong}[1]{#1\uparrow^\eta}
+\newcommand{\almostetalong}[1]{\overline{#1}}
+
+In addition, we are particularly interested in interpretation terms in
+\emph{$\eta$-long form}.
+
+\begin{definition}
+For a given variable environment $\Gamma$ and interpretation term $s$
+which is well-typed under $\Gamma$, the \emph{$\eta$-long form} of $s$,
+denoted $\etalong{s}$, is defined as follows:
+\begin{itemize}
+\item if $\Gamma \vdash s : \sigma \arrtype \tau$ and $s = \abs{x}{s'}$,
+  then $\etalong{s} =  \abs{x}{(\etalong{s'})}$
+\item if $\Gamma \vdash s : \sigma \arrtype \tau$ and $s$ does not have
+  the form $\abs{x}{s'}$, then $\etalong{s} =  \abs{x}{(
+  \etalong{(\app{s}{x})})}$
+\item if $\Gamma \vdash s : \quant{\alpha}{\sigma}$ and $s = \tabs{
+  \alpha}{s'}$, then $\etalong{s} = \tabs{\alpha}{(\etalong{s'})}$
+\item if $\Gamma \vdash s : \quant{\alpha}{\sigma}$ and $s$ does not have
+  the form $\tabs{\alpha}{s'}$, then $\etalong{s} = \tabs{\alpha}{
+  (\etalong{(\tapp{s}{\alpha})})}$
+\item if $\Gamma \vdash s : \nat$ or $\Gamma \vdash s : \alpha$ or
+  $\Gamma \vdash s : \sigma \times \tau$, then $\etalong{s} =
+  \almostetalong{s}$, where
+  \begin{itemize}
+  \item $\almostetalong{s} = s$ if $s$ is a natural number, variable or
+    function symbol
+  \item $\almostetalong{s} = \pair{\etalong{s_1}}{\etalong{s_2}}$ if
+    $s = \pair{s_1}{s_2}$
+  \item $\almostetalong{s} = \etalong{s}$ if $s$ has the form
+    $\abs{x}{s'}$ or $\tabs{\alpha}{s'}$
+  \item $\almostetalong{s} = \app{\almostetalong{s_1}}{(\etalong{s_2})}$
+    if $s = \app{s_1}{s_2}$
+  \item $\almostetalong{s} = \tapp{\almostetalong{s'}}{\tau}$
+    if $s = \tapp{s'}{\tau}$
+  \end{itemize}
+\end{itemize}
+A term $s$ is in \emph{$\eta$-long form} if $s = \etalong{s}$.
+\end{definition}
+
+It is clear that the $\eta$-long form of an interpretation term is
+unique; the following lemma shows that it is also always well-defined.
+
+\begin{lemma}
+For all interpretation terms $s$, the function $\etalong{s}$ is
+well-defined.
+\end{lemma}
+
+\begin{proof}
+Define here the \emph{depth} function on interpretation terms as
+follows:
+\begin{itemize}
+\item $\mathit{depth}(n) = \mathit{depth}(x) = \mathit{depth}(\mathtt{f})
+  = 0$
+\item $\mathit{depth}(\pair{s}{t}) = \max(\mathit{depth}(s),\mathit{
+  depth}(t)) + 1$
+\item $\mathit{depth}(\abs{x}{s}) = \mathit{depth}(\tabs{\alpha}{s}) =
+  \mathit{depth}(s) + 1$
+\item $\mathit{depth}(\app{s}{t}) = \max(\mathit{depth}(s),\mathit{
+  depth}(t)+1)$
+\item $\mathit{depth}(\tapp{s}{\tau}) = \mathit{depth}(s)$
+\end{itemize}
+(This definition is clearly well-defined as the term under consideration
+becomes smaller each step.)
+We can see that $\etalong{x}$ is well-defined for all variables by
+induction on their type.  This we use to see that $\etalong{(\apps{s}{
+x_1}{x_n})}$ is well-defined for $s$ a variable, natural number or
+function symbol by induction on its type.  Thus, well-definedness of
+$\etalong{s}$ for all $s$ with depth $0$ follows; also $\almostetalong{
+s} = s$ is well-defined for such $s$.
+For interpretation terms of depth $\geq 1$, well-definedness of
+$\etalong{s}$ and $\overline{s}$ follows by induction on
+$(\mathit{depth}(s)$, $1$, the size of the type of $s)$ in the case
+of $\etalong{s}$ and
+$(\mathit{depth}(s)$, $0$, the size of $s)$ in the case of
+$\almostetalong{s}$.
+\end{proof}
+
+\begin{lemma}
+If $s$ is a closed interpretation term in $\eta$-long form which is
+terminating under $\leadsto$, then there is a final interpretation term
+$t$ such that $s \leadsto^* t$.
+\end{lemma}
+
+\begin{proof}
+TODO
+\CK{And honestly not sure whether we even need this.}
+\end{proof}
+
+\LC{What do we need these $\eta$-expansions for? I think this just
+  works now with the modified $\leadsto$-rules for $\oplus,\otimes$.}
 

### 2018-11-28 museum.tex
more SN proof

@@
-% A museum of no-longer-needed lemmas and such (fragments of proofs of which may however become useful later)
+% A museum of no-longer-needed lemmas and such (fragments of proofs of
+% which may however become useful later)
+
+\begin{lemma}\label{lem_reducible_type}
+  If $\tau$ is a type and $t \in \val{\tau}{\xi}{\omega}$ then
+  $t : \omega(\tau)$.
+\end{lemma}
+
+\begin{proof}
+  Let~$\varphi$ be a type constructor. By induction on the kind
+  of~$\varphi$ we define when $\val{\varphi}{\omega}{\xi}$ is
+  \emph{good}:
+  \begin{itemize}
+  \item if $\varphi$ is a type then $\val{\varphi}{\xi}{\omega}$ is
+    good when $t \in \val{\varphi}{\xi}{\omega}$ implies
+    $t : \omega(\varphi)$,
+  \item if $\varphi$ has kind~$\kappa_1\arrkind\kappa_2$ then
+    $\val{\varphi}{\xi}{\omega}$ is good when
+    $\val{\varphi}{\xi[\subst{\alpha}{X}]}{\omega[\subst{\alpha}{\tau}]}$
+    is good for every $\tau \in \Tc_{\kappa_1}$ and every
+    $X \in \Cb_\tau$.
+  \end{itemize}
+  Now prove by induction on~$\varphi$
+  that~$\val{\varphi}{\omega,\xi}{}$ is good.
+
+  TODO
+\end{proof}
+
+Thus, we have:
+
+\begin{lemma}\label{lem_abusive_notation}
+  Every interpretation term has the form $s t_1 \dots t_n$ with $s$ a
+  variable, function symbol or abstraction $\abstraction{a}{s'}$ (for
+  $n \geq 0$), or $s$ a natural number and $n = 0$.
+\end{lemma}
+
+\begin{proof}
+By induction on the size of interpretation terms (and a simple case
+analysis).\qed
+\end{proof}
 

### 2018-12-28 museum.tex
restructuring

@@
 % which may however become useful later)
+
+$\{ \oplus_\sigma : \sigma \arrtype
+  \sigma \arrtype \sigma,\ \otimes_\sigma : \sigma \arrtype \sigma
+  \arrtype \sigma,\ \flatten_{\sigma} : \sigma \arrtype
+  \nat,\ \lift_{\sigma} : \nat \arrtype \sigma \mid \sigma \in \ITypes
+  \}$
+
+\section{TODO list}
+
+\renewcommand{\theenumii}{\arabic{enumi}.\arabic{enumii}}
+
+Here's what we need to do:
+\begin{enumerate}
+\item Define a world $\World$ and a well-founded ordering $\succ$ on
+  $\World$:
+  \begin{enumerate}
+  \item Define a set of terms $\World$ typed under some variation of
+    System F-$\omega$.
+  \item Define relations $\succ$ and $\succeq$ on the elements of $\World$.
+  \item Prove that $\succ$ is a well-founded ordering relation and that
+    $\succeq$ is a compatible quasi-ordering.
+  \end{enumerate}
+\item Specify what systems we are interested in analysing, and prove
+  standard results which will make their analysis doable.
+  \begin{enumerate}
+  \item Specify a form of system which includes all the systems of interest.
+  \item Specify a default way of interpreting terms in these systems.
+  \item Prove that in all such systems, using our way of interpreting
+    terms: if $\interpret{\ell} \succ \interpret{r}$ (resp.\ $\interpret{
+    \ell} \succeq \interpret{r}$) for a rule  $\ell \to r$, then
+    $\interpret{s} \succ \interpret{t}$ whenever $s \arr{\Rules} t$ by
+    this rule (resp.\ $\interpret{s} \succeq \interpret{t}$).
+  \end{enumerate}
+\item Obtain useful lemmas regarding these defaults.
+  \begin{enumerate}
+  \item $\interpret{s[x:=t]} = \interpret{s}[x:=\interpret{t}]$.
+  \item $\interpret{s\sigma} = \interpret{s}\sigma$.
+  \item \dots?
+  \end{enumerate}
+\item For some system of interest, prove its termination:
+  \begin{enumerate}
+  \item Present the system and give interpretations (following the
+    default scheme) for all ways of constructing terms.
+  \item Show that $\ell \succ r$ or $\ell \succeq r$ for all rules.
+    Remove the rules which are oriented using $\succ$ and repeat,
+    until all rules have been removed.
+  \end{enumerate}
+\end{enumerate}
+
+\renewcommand{\theenumii}{\alph{enumii}}
+
+We use \emph{rule removal}:
+
+\begin{theorem}\label{thm:ruleremove}
+Let $\Rules = \Rules_1 \cup \Rules_2$, and suppose that $\arr{\Rules_1}\:
+\subseteq\:\succ$ and $\arr{\Rules_2}\:\subseteq\:\succeq$ for a
+well-founded ordering $\succ$ and a compatible quasi-ordering $\succeq$.
+Then $\arr{\Rules}$ is terminating if and only if $\arr{\Rules_2}$ is
+(which is certainly the case if $\Rules_2 = \emptyset$).
+\end{theorem}
+
+\begin{proof}
+By well-foundedness of $\succ$, every infinite decreasing $\arr{\Rules}$
+sequence can only use finitely many steps using $\arr{\Rules_1}$.
+\qed
+\end{proof}
+
+This gives rise to the following algorithm:
+\begin{enumerate}
+\item While $\Rules$ is non-empty:
+  \begin{enumerate}
+  \item Orient all rules in $\Rules$ using $\succeq$ or $\succ$; at least
+    one of them must be oriented using $\succ$.
+  \item Remove all rules ordered by $\succ$ from $\Rules$.
+  \end{enumerate}
+\end{enumerate}
+If this algorithm succeeds, we have proven termination.
+
+\section{Lemmas}
 

### 2019-01-03 museum.tex
fixed succ

@@
 % which may however become useful later)
+
+  \begin{enumerate}
+  \item By coinduction we show $(s \oplus t) u_1 \ldots u_m
+    \succeq_\sigma s u_1 \ldots u_m$ for closed $u_1,\ldots,u_m$. The
+    second case is similar.
+
+    First note that $(s \oplus t) u_1 \ldots u_m \leadsto^* s u_1
+    \ldots u_m \oplus t u_1 \ldots u_m$.
+
+    If $\sigma = \nat$ then $((s \oplus t) u_1 \ldots u_m)\da = (s u_1
+    \ldots u_m)\da + (t u_1 \ldots u_m)\da \ge (s u_1 \ldots
+    u_m)\da$. Hence $(s \oplus t) u_1 \ldots u_m) \succeq_\nat s u_1
+    \ldots u_m$.
+
+    If $\sigma = \tau_1\arrtype\tau_2$ then by the coinductive
+    hypothesis $(s \oplus t) u_1 \ldots u_m q \succeq_{\tau_2} s u_1
+    \ldots u_m q$ for any $q \in \World_{\tau_1}$. Hence $(s \oplus t)
+    u_1 \ldots u_m \succeq_\sigma s u_1 \ldots u_m$.
+
+    If $\sigma = \forall(\alpha:\kappa)[\tau]$ then by the coinductive
+    hypothesis $(s \oplus t) u_1 \ldots u_m \xi \succeq_{\sigma'} s
+    u_1 \ldots u_m \xi$ for any closed $\xi \in \Tc_\kappa$, where
+    $\sigma' = \tau[\subst{\alpha}{\xi}]$. Hence $(s \oplus t) u_1
+    \ldots u_m \succeq_\sigma s u_1 \ldots u_m$.
+  \item By coinduction we show $(s \oplus (\lift n)) u_1 \ldots u_m
+    \succeq_\sigma s u_1 \ldots u_m$ for closed $u_1,\ldots,u_m$. The
+    second case is similar.
+
+    Note that $(s \oplus (\lift n)) u_1 \ldots u_m \leadsto^* s u_1
+    \ldots u_m \oplus n$. From this the base case $\sigma=\nat$
+    follows. The other cases follow from the induction hypothesis,
+    like in the first point above.
+  \end{enumerate}
 

### 2019-01-21 museum.tex
removed parametricity

@@
 % which may however become useful later)
+
+\subsection{Parametricity}
+
+To be filled in much later\ldots
+
+\begin{definition}\label{def_param_candidate}\normalfont
+  For two type constructors~$\tau_1,\tau_2$ of the same kind~$\kappa$,
+  we define by induction on~$\kappa$ the set~$\Rel_{\tau_1,\tau_2}$ of
+  $\tau_1,\tau_2$-candidates.
+
+  Assume $\kappa=*$, i.e., $\tau_1,\tau_2$ are types. A relation~$R$
+  between terms of type~$\tau_1$ and terms of type~$\tau_2$ is a
+  \emph{candidate relation} in $\Rel_{\tau_1,\tau_2}$ when the
+  following properties are satisfied:
+  \begin{enumerate}
+  \item if $R(t_1,t_2)$ and $t_1' \leadsto t_1$ (resp.~$t_2' \leadsto
+    t_2$) then $R(t_1',t_2)$ (resp.~$R(t_1,t_2')$),
+  \item \ldots
+  \end{enumerate}
+\end{definition}
+
+\begin{lemma}\label{lem_beta_param_candidate}
+  If $\sigma_1 =_\beta \sigma_1'$ and $\sigma_2 =_\beta \sigma_2'$
+  then $\Rel_{\sigma_1,\sigma_2} = \Rel_{\sigma_1',\sigma_2'}$.
+\end{lemma}
+
+\begin{proof}
+  Induction on the kind of~$\sigma_1,\sigma_2$.
+\end{proof}
+
+\begin{definition}\label{def_param_valuation}\normalfont
+  Let $\omega$ be a mapping from type constructor variables to pairs
+  of type constructors (respecting kinds). The mapping~$\omega$
+  extends in an obvious way to a mapping from type constructors to
+  pairs type constructors. We write $\omega_i(\tau)$ for the $i$-th
+  component of the pair~$\omega(\tau)$. A mapping~$\omega$ is
+  \emph{closed for~$\sigma$} if $\omega_i(\alpha)$ is closed for
+  $\alpha \in \FV(\sigma)$ (then $\omega_i(\sigma)$ is closed), for
+  $i=1,2$.
+
+  An \emph{$\omega$-valuation} is a mapping~$\xi$ on type constructor
+  variables such that $\xi(\alpha) \in
+  \Rel_{\omega_1(\alpha),\omega_2(\alpha)}$.
+
+  For each type constructor~$\sigma$, each mapping~$\omega$ closed
+  for~$\sigma$, and each $\omega$-valuation~$\xi$, we define
+  $\rval{\sigma}{\xi}{\omega}$ by induction on~$\sigma$:
+  \begin{itemize}
+  \item $\rval{\alpha}{\xi}{\omega} = \xi(\alpha)$ for a type
+    constructor variable~$\alpha$,
+  \item $\rval{\nat}{\xi}{\omega}(t_1,t_2)$ iff $t_1 \approx_\nat t_2$,
+  \item $\rval{\sigma \arrtype \tau}{\xi}{\omega}(t_1,t_2)$ iff $t_i :
+    \omega_i(\sigma\arrtype\tau)$ and for all $s_1,s_2$ such that
+    $\rval{\sigma}{\xi}{\omega}(s_1,s_2)$ we have
+    $\rval{\tau}{\xi}{\omega}(t_1 s_1, t_2 s_2)$,
+  \item $\rval{\forall(\alpha:\kappa)[\sigma]}{\xi}{\omega}(t_1,t_2)$
+    iff $t_i : \omega_i(\forall\alpha\sigma)$ and for all closed type
+    constructors~$\tau_1,\tau_2$ of kind~$\kappa$ and every $R \in
+    \Rel_{\tau_1,\tau_2}$ we have
+    $\rval{\sigma}{\xi[\subst{\alpha}{R}]}{\omega[\subst{\alpha}{(\tau_1,\tau_2)}]}$.
+  \end{itemize}
+\end{definition}
 

### 2019-02-05 museum.tex
abstract
comments and changes

@@
 
+Polymorphism in type theory is parametric, which means that a
+polymorphic function behaves uniformly with respect to its generic
+arguments, not depending on their actual types. This makes it
+impossible to directly extract elements from a heterogenous list -- we
+can only perform some operations on them under the assumption that we
+know nothing about their types. In practice, one would constrain the
+type variable~$\alpha$ above with a type class, in order to guarantee
+the existence of certain functions on the elements of a list.
+
 \subsection{Parametricity}
-
-To be filled in much later\ldots
 

### 2019-02-11 museum.tex
changes and comments
final version

@@
 % which may however become useful later)
+
+\cite[Chapter 6,7]{Kop2012}).
+\CK{If space happens to permit, we can replace the pointer to my thesis by
+cite\{kop:raa:12,suz:kus:bla:11\}, but this does require 4 extra lines
+of references.}
+
+From the definition we immediately see that all encodings have the expected types:
+
+\begin{lemma}\label{lem:encodings_types}
+  \begin{enumerate}
+  \item If $\Gamma \proves t_1 : \sigma$ and $\Gamma \proves t_2 :
+    \tau$ then $\Gamma \proves \pair{t_1}{t_2} : \sigma \times \tau$.
+  \item If $\Gamma \proves t : \tau_1 \times \tau_2$ then $\Gamma
+    \proves \pi^i(t) : \tau_i$.
+  \item If $\Gamma \proves t : \sigma[\subst{\alpha}{\tau}]$ then
+    $\Gamma \proves \expair{\tau}{t} : \Sigma \alpha . \sigma$.
+  \item If $\Gamma \proves t : \Sigma \alpha . \sigma$ and
+    $\Gamma,x:\sigma \proves s : \rho$ and $\alpha \notin
+    \FTV(\Gamma,\rho)$ then $\xlet{}{t}{\alpha,x}{s} : \rho$.
+  \end{enumerate}
+\end{lemma}
+
+In addition, $\beta$-reduction provides the desired reduction rules:
+
+\begin{lemma}\label{lem:encodings_reduce}
+  $\pi^i(\pair{t_1}{t_2}) \leadsto_\beta^* t_i$ and
+  $\xlet{}{\expair{\tau}{t}}{\alpha,x}{s} \leadsto_\beta^*
+  s[\subst{\alpha}{\tau}][\subst{x}{t}]$.
+\end{lemma}
 

### 2019-03-28 rebuttal.txt
Starting the rebuttal.

We thank the reviewers for their thorough reading of the article and 
for their valuable feedback.  We shall apply the reviewers' comments
to improve the article for the final version.

On the reviewers' main concerns:

R1: TODO

R2: TODO

R3 notes that we do not define a dependency pair notion or any automation.

These two points are currently named as lines for future work.  We believe
that it would not benefit the present work to do too much at once.
Automation is challenging because it requires the development of several new
ideas; setting up a dependency pair approach would be a full, mostly separate
line of research.  (However, it is worth noting that a future definition of
dependency pairs is likely to benefit greatly from the existence of a
reduction pair, like the one we define here).

### 2019-03-29 rebuttal.txt
more rebuttal

@@
 
-R1: TODO
+R1 additional references
 
-R2: TODO
+Thank you for making us aware of more related work; we will include
+the references.
+
+R2  why F-omega
+
+In F-omega there are type constructors, which we need to express
+e.g. type application A (l.177) and existentials (let, ext). Having
+type constructors in the system makes the whole presentation more
+uniform with little cost to the complexity of the proofs.
+
+R2 concern about the J tau tau counterexample
+
+The termination proof method for the system in section 4.1 is indeed
+standard, but the question itself is quite subtle, as briefly
+mentioned in lines 230-239. The crucial property of system F-omega is
+parametricity -- that functions behave "uniformly" with respect to
+their type arguments, e.g., a function (\alpha : * . t) cannot check
+if alpha is equal to some other type or inspect it in any way. When an
+extension breaks this property, this often results in
+non-termination. The extension with J tau tau -> \x : tau . x breaks
+parametricity because the rule being non-left-linear it can
+effectively compare two arbitrary types, including type
+variables. This cannot be defined in our system; \tau : * . \x : tau
+. lift tau (flatten tau x) seems like such a definition, but it's not
+-- it doesn't work for arbitrary types, only for tau=nat; it gets
+"stuck" on type variables, and is not an identify for
+tau/=nat. Moreover, the natural numbers cannot be distinguished
+"inside" the system. In other words, we could replace all natural
+number constants with 0 and this wouldn't change the reduction
+behaviour of terms. So for the purposes of termination, the type nat
+is essentially a singleton. This implies that, while we have
+polymorphic functions between an arbitrary type alpha and nat which
+are not constant when seen "from outside" the system, they are
+constant for the purposes of reduction "inside" the system (as they
+would have to be in a parametric system). Intuitively, these
+properties of our system ensure that it stays "close enough" to
+F-omega so that the standard termination proof still generalises.
+
+R1/R2 question about coinduction
+
+A definition by coinduction simply seemed most natural to the
+authors. Lemma 4.9 stating a non-coinductive reformulation of this
+definition was added as an afterthought. The non-coinductive
+reformulation of the proofs is straightforward enough that we could do
+it for the final version of the paper, if the reviewers so
+desire. Only the "easy" proofs of basic properties of > use
+coinduction, later the details of the definition don't matter, only
+that the lemmas hold. One advantage of coinduction in this setting,
+and this ties in with the question about formalisation, is that the
+current definition is more precise in a sense. The coinductive proofs
+could be typed into a proof assistant almost as they stand. When
+reformulated in an alternative manner, they would be harder to
+formalise. In fact, the easiest way to formalise them would probably
+be to reformulate them back in a coinductive way.
+
+R1 l273. Shouldn't the domain of C be the free variables of t and s?
+
+This is equivalent, because the value of C on variables not free in
+t,s doesn't matter.
+
+R2 avoid introducing the constants flatten, ..., and instead DEFINE those with induction
+
+Induction doesn't work because of impredicative polymorphism. We can
+have e.g. "\alpha : * . flatten alpha t". Then how to define "flatten
+alpha t" by induction when alpha is a type variable? Note that when
+applied to an arbitrary type tau this will reduce to "flatten tau t".
+
+R2 is "s>t" computable?
+
+Not in general, I don't think so.
+
+R2 l.435
+
+cons has type forall alpha . alpha -> List -> List, so only the first
+three lambda-abstractions count as arguments, the others are just the
+details of the encoding of lists in our system.
+
+R2 why not add 2^x to the basic functions
+
+This could be done without any fundamental problems, but wouldn't help
+with interpreting the few remaining rules, where the problem is the
+limitation of the current encoding of impredicative existential
+quantification, which makes certain terms incomparable.
 

### 2019-03-29 rebuttal.txt
minor

@@
 non-termination. The extension with J tau tau -> \x : tau . x breaks
-parametricity because the rule being non-left-linear it can
+parametricity because the rule, being non-left-linear, can
 effectively compare two arbitrary types, including type
@@
 -- it doesn't work for arbitrary types, only for tau=nat; it gets
-"stuck" on type variables, and is not an identify for
+"stuck" on type variables, and is not an identity for
 tau/=nat. Moreover, the natural numbers cannot be distinguished
@@
 cons has type forall alpha . alpha -> List -> List, so only the first
-three lambda-abstractions count as arguments, the others are just the
-details of the encoding of lists in our system.
+two non-type lambda-abstractions count as arguments, the others are
+just the details of the encoding of lists in our system.
 

### 2019-03-31 rebuttal.txt
Some updates.

@@
 
-R1 additional references
+* R1 notes some additional references that we did not include.
 
 Thank you for making us aware of more related work; we will include
-the references.
+these references.
 
-R2  why F-omega
+* R1/R2 question why we use coinduction, which not everyone is
+familiar with.
+
+A definition by coinduction simply seemed most natural to the
+authors. Lemma 4.9, which states a non-coinductive reformulation of
+this definition, was added as an afterthought. The non-coinductive
+reformulation of the proofs is straightforward enough that we could do
+it for the final version of the paper, if the reviewers so desire.
+Only the "easy" proofs of basic properties of > use coinduction,
+later the details of the definition don't matter, only
+that the lemmas hold.
+
+One advantage of coinduction in this setting -- and this ties in with
+the question about formalisation -- is that the current definition is,
+in a sense, more precise. The coinductive proofs could be typed into a
+proof assistant almost as they stand. When reformulated in an
+alternative manner, they would be harder to formalise. (And in fact,
+the easiest way to formalise them would probably be to reformulate
+them back in a coinductive way.)
+
+>>> CK: added this one! <<<
+
+* R1 notes that we do not discuss differences with v.d. Pol's work.
+
+It is true that our definitions are, in a sense, less general than
+those of [8]; this work should be seen more as an extension of [1]
+(which restricts the ideas of [8] to a simpler class). We will make
+this more clear in the final version.
+
+* R1 questions whether in l273, the domain of C should be restricted
+to the free variables of t and s.
+
+This is equivalent, because the value of C on variables not free in
+t,s doesn't matter.
+
+* R2 wonders why we use F-omega rather than F.
 
@@
 
-R2 concern about the J tau tau counterexample
+* R2 expresses concern about the J tau tau counterexample.
 
@@
 extension breaks this property, this often results in
-non-termination. The extension with J tau tau -> \x : tau . x breaks
-parametricity because the rule, being non-left-linear, can
-effectively compare two arbitrary types, including type
-variables. This cannot be defined in our system; \tau : * . \x : tau
-. lift tau (flatten tau x) seems like such a definition, but it's not
--- it doesn't work for arbitrary types, only for tau=nat; it gets
-"stuck" on type variables, and is not an identity for
-tau/=nat. Moreover, the natural numbers cannot be distinguished
-"inside" the system. In other words, we could replace all natural
-number constants with 0 and this wouldn't change the reduction
-behaviour of terms. So for the purposes of termination, the type nat
-is essentially a singleton. This implies that, while we have
+non-termination.
+
+The extension with J tau tau -> \x : tau . x breaks parametricity
+because the rule, being non-left-linear, can effectively compare two
+arbitrary types, including type variables. This cannot be defined in
+our system; \tau : * . \x : tau. lift tau (flatten tau x) seems like
+such a definition, but it's not -- it doesn't work for arbitrary
+types, only for tau=nat; it gets "stuck" on type variables, and is not
+an identity for tau/=nat. Moreover, the natural numbers cannot be
+distinguished "inside" the system. In other words, we could replace
+all natural number constants with 0 and this wouldn't change the
+reduction behaviour of terms. So for the purposes of termination, the
+type nat is essentially a singleton. This implies that, while we have
 polymorphic functions between an arbitrary type alpha and nat which
@@
 
-R1/R2 question about coinduction
-
-A definition by coinduction simply seemed most natural to the
-authors. Lemma 4.9 stating a non-coinductive reformulation of this
-definition was added as an afterthought. The non-coinductive
-reformulation of the proofs is straightforward enough that we could do
-it for the final version of the paper, if the reviewers so
-desire. Only the "easy" proofs of basic properties of > use
-coinduction, later the details of the definition don't matter, only
-that the lemmas hold. One advantage of coinduction in this setting,
-and this ties in with the question about formalisation, is that the
-current definition is more precise in a sense. The coinductive proofs
-could be typed into a proof assistant almost as they stand. When
-reformulated in an alternative manner, they would be harder to
-formalise. In fact, the easiest way to formalise them would probably
-be to reformulate them back in a coinductive way.
-
-R1 l273. Shouldn't the domain of C be the free variables of t and s?
-
-This is equivalent, because the value of C on variables not free in
-t,s doesn't matter.
-
-R2 avoid introducing the constants flatten, ..., and instead DEFINE those with induction
+* R2 queries whether we cannot avoid introducing the constants flatten,
+..., and instead DEFINE those with induction.
 
 Induction doesn't work because of impredicative polymorphism. We can
-have e.g. "\alpha : * . flatten alpha t". Then how to define "flatten
-alpha t" by induction when alpha is a type variable? Note that when
-applied to an arbitrary type tau this will reduce to "flatten tau t".
+have, e.g., "\alpha : * . flatten alpha t". But then how to define
+"flatten alpha t" by induction when alpha is a type variable? Note
+that when applied to an arbitrary type tau this will reduce to
+"flatten tau t".
 
-R2 is "s>t" computable?
+>>> CK: I don't actually understand this question, so I don't know how best to rephrase it in the form "R2 asks" or whatnot. <<
+
+* R2 is "s>t" computable?
 
@@
 
-R2 l.435
+* R2 asks for a better explanation of l.435.
 
-cons has type forall alpha . alpha -> List -> List, so only the first
-two non-type lambda-abstractions count as arguments, the others are
-just the details of the encoding of lists in our system.
+Since cons has type forall alpha . alpha -> List -> List, only the
+first two non-type lambda-abstractions count as arguments; the others
+are just the details of the encoding of lists in our system.
 
-R2 why not add 2^x to the basic functions
+* R2 questions whether we could add 2^x to the basic functions, and
+whether it would be helpful to complete our "larger example".
 
 This could be done without any fundamental problems, but wouldn't help
-with interpreting the few remaining rules, where the problem is the
+with interpreting the few remaining rules: there, the problem is the
 limitation of the current encoding of impredicative existential
@@
 
-R3 notes that we do not define a dependency pair notion or any automation.
+* R3 notes that we do not define a dependency pair notion or any
+automation.
 
-These two points are currently named as lines for future work.  We believe
-that it would not benefit the present work to do too much at once.
-Automation is challenging because it requires the development of several new
-ideas; setting up a dependency pair approach would be a full, mostly separate
-line of research.  (However, it is worth noting that a future definition of
-dependency pairs is likely to benefit greatly from the existence of a
-reduction pair, like the one we define here).
+These two points are currently named as lines for future work.  We
+believe that it would not benefit the present work to do too much at
+once. Automation is challenging because it requires the development of
+several new ideas; setting up a dependency pair approach would be a
+full, mostly separate line of research.
+
+(However, it is worth noting that a future definition of dependency
+pairs is likely to benefit greatly from the existence of a reduction
+pair, like the one we define here).
+
+CK: Things we have not discussed so far:
+
+* R1 observes that we do not discuss the decidability of our main
+ordering.
+
+CK: no-op on this one? We did respond to the comments about
+co-induction.
+
+* R1 wonders asks us to put an intuition of the proof for Theorem 4.4
+in the paper's body, and questions whether some termination criterion
+could be applied here.
+
+CK: the first is simple ("sure, if space permits") and probably
+doesn't need to be explicitly addressed (we have already promised to
+improve things before the final version, which will include the
+presentation). For the second, I don't know what the reviewer means...

### 2019-04-01 rebuttal.txt
updates

@@
 >>> CK: I don't actually understand this question, so I don't know how best to rephrase it in the form "R2 asks" or whatnot. <<
+>>> LC: I reformulated it the way I interpret it.
 
-* R2 is "s>t" computable?
+* R2 asks if the following problem is decidable: given two interpretation terms s,t, does s>t hold?
 
-Not in general, I don't think so.
+Not in general, I don't think so. However, one can come up with
+calculation rules like in Lemmas 6.3-6.5 to automatically handle
+common cases in practice.
 
@@
 
-* R2 questions whether we could add 2^x to the basic functions, and
+* R2 asks whether we could add 2^x to the basic functions, and
 whether it would be helpful to complete our "larger example".
@@
 
-* R1 wonders asks us to put an intuition of the proof for Theorem 4.4
+LC: I think this is the s>t question
+
+* R1 asks us to put an intuition of the proof for Theorem 4.4
 in the paper's body, and questions whether some termination criterion
@@
 presentation). For the second, I don't know what the reviewer means...
+
+LC: Termination criterion for that one? I don't think so -- it
+crucially depends on impredicative polymorphism, and we're just
+introducing methods for it!

### 2019-04-01 rebuttal.txt
Changes approved.

@@
 them back in a coinductive way.)
-
->>> CK: added this one! <<<
 
@@
 
->>> CK: I don't actually understand this question, so I don't know how best to rephrase it in the form "R2 asks" or whatnot. <<
->>> LC: I reformulated it the way I interpret it.
+* R2 asks if the following problem is decidable: given two
+interpretation terms s,t, does s>t hold?
 
-* R2 asks if the following problem is decidable: given two interpretation terms s,t, does s>t hold?
-
-Not in general, I don't think so. However, one can come up with
+Not in general, we believe. However, one can come up with
 calculation rules like in Lemmas 6.3-6.5 to automatically handle
@@
 pair, like the one we define here).
-
-CK: Things we have not discussed so far:
-
-* R1 observes that we do not discuss the decidability of our main
-ordering.
-
-CK: no-op on this one? We did respond to the comments about
-co-induction.
-
-LC: I think this is the s>t question
-
-* R1 asks us to put an intuition of the proof for Theorem 4.4
-in the paper's body, and questions whether some termination criterion
-could be applied here.
-
-CK: the first is simple ("sure, if space permits") and probably
-doesn't need to be explicitly addressed (we have already promised to
-improve things before the final version, which will include the
-presentation). For the second, I don't know what the reviewer means...
-
-LC: Termination criterion for that one? I don't think so -- it
-crucially depends on impredicative polymorphism, and we're just
-introducing methods for it!
